Characterizing subarctic biomes for land surface modeling of pollution and climate risk

### 2021-04-10 VegOzone.tex
Last comment so far by Ane

\subsubsection{Temperature, precipitation, and global irradiance}
\label{subsubsec:anomal_tpq}
In Fig.~\ref{fig:anomalies_svanvik}a), temperature anomalies at Svanvik are shown as percentage of days warmer/colder than $\pm 1\,\sigma$ from the climatological mean for each month. Negative deviations from climatology are displayed as negative percentage. The full annual positive/negative deviations are indicated in the respective corners (right upper/right lower). The hatched area between the dashed lines indicates the expected percentage of values falling above/below $\pm 1\,\sigma$ if a normal distribution is assumed ($15.9\,\unit{\%}$). The summer of 2018 was significantly warmer than average. Especially in May and July more than $40\,\unit{\%}$ of days were significantly warmer than normal temperature. Temperatures continued to be significantly warmer throughout fall into early winter (August--November) with the exception of October. March 2018 was unusually cold. The summer of 2019 was fairly average, however significantly colder in July, while August was both significantly colder and warmer. Temperatures in April were significantly warmer than average.

\begin{figure}[t]
  \includegraphics[width=12cm]{fig08}
  \caption{2018/19 anomalies of key environmental variables at Svanhovd displayed as percentages of days deviating from climatological mean by means of standard deviation for each month. Negative deviations from climatology are shown as negative percentage. The annual positive/negative deviations are indicated in the respective corners (right upper/right lower). The hatched area between the dashed lines indicates statistical significance under the assumption of a normal distribution ($15.9\,\unit{\%}$, $30.8\,\unit{\%}$ for Precip.). (a) Temperature, days warmer/colder than $\pm 1\,\sigma$; (b) precipitation, days wetter/drier than $\pm \frac{1}{2}\,\sigma$; (c) global irradiance, days lighter/darker than $\pm 1\,\sigma$; (d) ozone, days higher/lower than $\pm 1\,\sigma$. The July 2018 ozone data corrected for data acquisition problems is indicated with the star (refer Section~\ref{subsec:ozone_reco} for details). The growing season of 2018 was significantly warmer, brighter, and slightly drier than average, especially in May and July. Ozone was significantly elevated in May. The growing season of 2019 was fairly average but significantly warmer than average in April and colder in July/August. No significant elevation in ozone was found.}
  \label{fig:anomalies_svanvik}
\end{figure}

%\begin{figure}[t]
  %\includegraphics[width=8.3cm]{temperature_signific_1}
  %\includegraphics[width=8.3cm]{fig08}
  %\caption{Temperature anomalies at Svanvik. Percentage of days warmer/colder than $\pm 1\,\sigma$ from climatological mean displayed for each month. Negative deviations from climatology are shown as negative percentage. The annual positive/negative deviations are indicated in the respective corners (right upper/right lower). The hatched area between the dashed lines indicates the expected percentage of values falling above/below $\pm 1\,\sigma$ if a normal distribution is assumed ($15.9\,\unit{\%}$). The summer of 2018 was significantly warmer than average, especially May and July. The summer of 2019 was fairly average but significantly warmer than average in April and colder in July/August.}
  %\label{fig:plot_temperature_anomalies_svanvik}
%\end{figure}

In a similar way, the precipitation anomalies at Svanvik are displayed in Fig.~\ref{fig:anomalies_svanvik}b). However, the percentage of days wetter/drier than $\pm \frac{1}{2}\,\sigma$ from the climatological mean is shown for each month, due to the larger variability in observed precipitation too few days showed a significance of $1 \sigma$ or higher. As above, negative deviations from the climatology are shown as negative percentage and annual positive/negative deviations are indicated in the respective corners. The hatched area between the dashed lines indicates the expected percentage of values falling above $\pm\frac{1}{2}\,\sigma$ if a normal distribution is assumed ($30.8\,\unit{\%}$). Unlike for the temperature anomalies, the picture is not as clear. While March 2018 was unusually wet, summer and fall (May/July, September/October) were significantly drier than average. 2019 was rather average throughout the year, with drier than usual fall (September/October) and a wet December.

%\begin{figure}[t]
  %\includegraphics[width=8.3cm]{precipitation_signific_1}
  %\includegraphics[width=8.3cm]{fig09}
  %\caption{Precipitation anomalies at Svanvik. Percentage of days wetter/drier than $\pm \frac{1}{2}\,\sigma$ from climatological mean displayed for each month. Negative deviations from climatology are shown as negative percentage. The annual positive/negative deviations are indicated in the respective corners (right upper/right lower). The hatched area between the dashed lines indicates the expected percentage of values falling above $\pm\frac{1}{2}\,\sigma$ if a normal distribution is assumed ($30.8\,\unit{\%}$). Summer was significantly dry only in 2018, while fall was significantly dry in both years.}
  %\label{fig:plot_precipitation_anomalies_svanvik}
%\end{figure}

Similarly, the global irradiance anomalies at Svanvik are presented in Fig.~\ref{fig:anomalies_svanvik}c). The percentage of days above/below $\pm 1\,\sigma$ from the climatological mean is displayed for each month. The hatched area between the dashed lines indicates the expected percentage of values falling above/below $\pm 1\,\sigma$ if a normal distribution is assumed. In summer 2018 (May/July), global irradiance was significantly higher than on average, while 2019 was generally an average year with a slightly darker spring (April/May).

%\begin{figure}[t]
  %\includegraphics[width=8.3cm]{global_rad_signific_1}
  %\includegraphics[width=8.3cm]{fig10}
  %\caption{Global irradiance anomalies at Svanvik. Percentage of days above/below $\pm 1\,\sigma$ from climatological mean displayed for each month. Negative deviations from climatology are shown as negative percentage. The annual positive/negative deviations are indicated in the respective corners (right upper/right lower). The hatched area between the dashed lines indicates the expected percentage of values falling above/below $\pm 1\,\sigma$ if a normal distribution is assumed ($15.9\,\unit{\%}$). In summer 2018 (May/July), global irradiance was significantly higher than on average, while 2019 had significantly less than normal irradiance in April/May and more in June.}
  %\label{fig:global_rad_signific}
%\end{figure}

In summary, 2018 was significantly warmer, drier, and lighter than the climatological average, while 2019 was a rather average year.

@@

If the data collected at Svanvik had been complete in 2018 and the climatology had extended into the resent past, we could have assessed the significance of ozone anomalies in the same way as for temperature, precipitation, and global irradiance. In Section~\ref{subsubsec:clim_ozone}, we described the construction of climatologies for northern Fennoscandia based on long-term observations at Esrange, Jergul/Karasjok, and Pallas. For Svanvik this assessment is based on rather uncertain observations from 1986--1996. We also discussed the general patterns, similarities and differences. We shall now investigate \chem{\Delta[O_3]} in 2018/19 with respect to these climatologies in more detail.

>>>>

\subsubsection{Temperature, precipitation, and global irradiance}
\label{subsubsec:anomal_tpq}
In Fig.~\ref{fig:anomalies_svanvik}a), temperature anomalies at Svanvik are shown as percentage of days warmer or colder than $\pm 1\,\sigma$ from the climatological mean for each month. Negative deviations from climatology are displayed as negative percentage. The full annual positive/negative deviations are indicated in the respective corners (right upper/right lower). The hatched area between the dashed lines indicates the expected percentage of values falling above/below $\pm 1\,\sigma$ if a normal distribution is assumed ($15.9\,\unit{\%}$). The summer of 2018 was significantly warmer than average. Especially in May and July more than $40\,\unit{\%}$ of the days were significantly warmer. Significantly warmer days continued to occure in all months (August--November) with the exception of October. March 2018 had many unusually cold days. The summer of 2019 was fairly average, however significantly colder days in July occured, while August had both significantly colder and warmer days. April 2019 had significantly warmer than average days.

\begin{figure}[t]
  \includegraphics[width=12cm]{fig08}
  \caption{2018/19 anomalies of key environmental variables at Svanhovd displayed as percentages of days significantly deviating from climatological mean by means of standard deviation for each month. Negative deviations from climatology are shown as negative percentage. The annual positive/negative deviations are indicated in the respective corners (right upper/right lower). The hatched area between the dashed lines indicates statistical significance under the assumption of a normal distribution ($15.9\,\unit{\%}$, $30.8\,\unit{\%}$ for Precip.). (a) Temperature; (b) precipitation; (c) global irradiance; (d) ozone.}
  \label{fig:anomalies_svanvik}
\end{figure}

In a similar way, the precipitation anomalies at Svanvik are displayed in Fig.~\ref{fig:anomalies_svanvik}b). Due to a large interannual variability in observed precipitation, only a few days deviated significantly from the climatology on the $1 \sigma$ level. Therefore, we lowered the $\sigma$ level constraint. The percentage of days wetter/drier than $\pm \frac{1}{2}\,\sigma$ from the climatological mean is shown for each month. Negative deviations from the climatology are shown as negative percentage and annual positive/negative deviations are indicated in the respective corners. The hatched area between the dashed lines indicates the expected percentage of values falling above $\pm\frac{1}{2}\,\sigma$ if a normal distribution is assumed ($30.8\,\unit{\%}$). Unlike for the temperature anomalies, the picture is not as clear. While March 2018 had a significant number of days wetter than unusual, summer and fall (May/July, September/October) had a significantly number of days drier than average. Overall, 2019 was rather average throughout the year, but with a significant number of days  drier than normal (September/October) and a wetter December.

Similarly, the global irradiance anomalies at Svanvik are presented in Fig.~\ref{fig:anomalies_svanvik}c). The percentage of days above or below $\pm 1\,\sigma$ from the climatological mean is displayed for each month. The hatched area between the dashed lines indicates the expected percentage of values falling above/below $\pm 1\,\sigma$ if a normal distribution is assumed. In summer 2018 (May/July), global irradiance showed a significant number of days with higher than average irradiance, while 2019 was generally an average year with a significant number of days darker in spring (April/May).

In summary, 2018 had a significant number of days which were warmer, drier, and brighter than the climatological average, while 2019 was a rather average year.

@@

If the data collected at Svanvik had been complete in 2018 and the climatology had extended into the recent past, we could have assessed the significance of ozone anomalies in the same way as for temperature, precipitation, and global irradiance. In Section~\ref{subsubsec:clim_ozone}, we described the construction of climatologies for northern Fennoscandia based on long-term observations at Esrange, Jergul/Karasjok, and Pallas. For Svanvik this assessment is based on rather uncertain observations from 1986--1996. We also discussed the general patterns, similarities and differences. We shall now investigate \chem{\Delta[O_3]} in 2018/19 with respect to these climatologies in more detail.
